Large Scale Structure Reconstruction with Short-Wavelength Modes

### 2019-10-09 paper/v1.tex
1

\date{\today}
\begin{abstract}
\noindent Large scale density modes cannot be measured directly through surveys but we can study them by measuring perturbations on small scales. Cosmological perturbation theory predicts that second-order density inhomogeneities are a convolution of a short- and a long-wavelength mode. This arises physically because small structure structure varies depending on the large scale environment in which it resides. This will induce small scale clustering in real space and an off-diagonal term in two-point statistics in Fourier space. Therefore we can build a quadratic estimator for the large scale field. We demonstrate that this quadratic estimator works amazingly well for our N-body simulation test.

\end{abstract}
\maketitle
\section{Introduction}
Measuring the distribution of structures on large scales has been one of the main goals of recent surveys despite it is hard to accomplish directly. To tackle this problem, relations between long- and short-wavelength modes have been frequently discussed recent years \cite{Baldauf:2011fer}\cite{Barreira:2017res}. Method of using small scale position-dependent power spectrum to compute squeezed-limit bispectrum also indicates that we can use small scale perturbation to infer large scale information to some extent \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to achieve this goal, we develop in this work a generalized approach of building a quadratic estimator to measure long-wavelength modes indirectly. \\
Starting from second-order standard perturbation theory (SPT) \cite{Goroff:1986sts}\cite{Makino:1992fs}\cite{Jain:1994sop}, we can identify second-order perturbation of a short-wavelength mode as a convolution of a short- and a long-wavelength mode. This can be further related to cosmic microwave background (CMB) lensing case \cite{Hu:2002mr}, where CMB temperature field has a second-order correction due to gravitational field along line of sight. Similarly in this case the short-wavelength mode's nonlinear terms are related to the large scale environment it belongs to.
\\
During CMB lensing quadratic estimator's construction, small scale two-point correlation of CMB temperature modes has off-diagonal terms due to large scale perturbations caused by gravitational lensing. The same procedure applies in our case as well - off-diagonal terms of small scale correlations are no-longer zero due to large scale modes. Thus we can create a quadratic estimator for long-wavelength modes using the exact same formalism. \\
We begin with a brief review of SPT up to second-order in section \ref{sec1}. Then we build the quadratic estimator and compute the detectability of the estimator in section \ref{sec2}. In section \ref{sec3} we present a concise study of the estimator with data from N-body simulation. Finally we conclude this work and discuss future work related to this in section \ref{sec4}. We use flat $\Lambda$CDM model with Planck Collaboration XVI (2014) \cite{Planck:2014cos} cosmological parameters in this work (same as parameters of the N-body simulation).

@@

In this section we review the SPT approach which describes small scale nonlinearity. Then we compute the off-diagonal term of two short-wavelength modes in the squeezed limit. This off-diagonal term proportional to a long-wavelength mode will be the key factor of constructing the quadratic estimator. \\

@@

Plugging the perturbative series \ec{pert1} and \ec{pert2} into the fluid equations \ec{fluid1}-\ec{fluid3} after Fourier transform, we can solve second-order density contrast and the result is given by:

@@

Notice \ec{f2} is derived in Einstein-de Sitter universe. Nonetheless when we actually do relative calculation in $\Lambda$CDM universe, the difference is negligible \cite{Takahashi:2008to} thus we still use the expression from \ec{f2} throughout the whole work.\\

@@

%Taking the divergence of \ec{eulereq} we can eliminate the gravitational potential $\Phi$ in Fourier space and get evolution equations for $\delta(\vk,\tau)$ and $\theta(\vk,\tau)$ in Fourier space:
%\bea 
%\frac{\partial{\delta}}{\partial\tau} + {\theta} &=& -\int \frac{d^{3}\vk_{1}}{(2\pi)^3} \int \frac{d^{3}\vk_{2}}{(2\pi)^3} (2\pi)^3 \delta_{\rm D}(\vk-\vk_{12}) \vs
%&& \times\frac{\vk \cdot \vk_1}{k_1^2}{\theta}(\vk_1,\tau){\delta}(\vk_2,\tau) \eql{fouriercont}\\
%\frac{\partial{\delta}}{\partial\tau}+\frac{da}{d\tau}\frac{{\theta}}{a}+\frac{6}{\tau^2}{\delta}&=&-\int \frac{d^{3}\vk_{1}}{(2\pi)^3} \int \frac{d^{3}\vk_{2}}{(2\pi)^3} (2\pi)^3 \delta_{\rm D}(\vk-\vk_{12})\vs
%&&\times \frac{k^2(\vk_1 \cdot \vk_2)}{2k_1^2 k_2^2} {\theta}(\vk_1,\tau){\theta}(\vk_2,\tau) \eql{fouriereuler}
%\eea
%here $\vk_{12}=\vk_1 +\vk_2$ and more generally $\vk_{1\cdots n}=\vk_1 + \cdots +\vk_n$; $\theta\equiv \vec{\nabla}\cdot \vec{v}$ is the divergence of velocity field.\\
%In Einstein-de Sitter space, linear growth function $D_{1}(a)=a$ and we can solve these equations perturbatively using expansion:
%\bea
%{\delta}(\vk,\tau)&=&\sum_{n=1}^{\infty} {\delta}^{(n)}(\vk,\tau)=\sum_{n=1}^{\infty}a^{n}(\tau)\delta_{n}(\vk) \\ {\theta}(\vk,\tau)&=&\sum_{n=1}^{\infty}{\theta}^{(n)}(\vk,\tau)=-H(\tau)%\sum_{n=1}^{\infty}a^{n+1}(\tau)\theta_{n}(\vk) 
%\eea
%\\where the superscript $(n)$ means the order of perturbation theory, and first order term ${\delta}^{(1)}$ corresponds to linear evolution. Linear power spectrum is given by this first order term via:
%\be 
%\langle {\delta}(\vk){\delta}(\vk') \rangle =(2\pi)^3 \delta_{\rm D}(\vk+\vk')P_{\rm lin}(k) \eql{lin}
%\ee 
%Substituting the two perturbative series into \ec{fouriercont} and \ec{fouriereuler} we get recursion relations for ${\delta}_{n}(\vk)$ and ${\theta}_{n}(\vk)$ with solution:
%\bea
%{\delta}_{n}(\vk) &=& \int \frac{d^{3}\vk_{1}}{(2\pi)^3} \cdots \int \frac{d^{3}\vk_{n}}{(2\pi)^3}(2\pi)^3 \delta_{\rm D} (\vk-\vk_{1\cdots n}) \vs
%&&\times F_{n}(\vk_1,\cdots,\vk_n){\delta}_1(\vk_{1}) \cdots {\delta}_1(\vk_{n}) \\
%{\theta}_{n}(\vk) &=& \int \frac{d^{3}\vk_{1}}{(2\pi)^3} \cdots \int \frac{d^{3}\vk_{n}}{(2\pi)^3}(2\pi)^3 \delta_{\rm D} (\vk-\vk_{1\cdots n}) \vs
%&&\times G_{n}(\vk_1,\cdots,\vk_n){\delta}_1(\vk_{1}) \cdots {\delta}_1(\vk_{n})
%\eea
%and recursion relations are encoded in kernels $F_n$ and $G_n$. Since we only consider up to second-order, the second-order symmetrized kernels are given by:
%\bea 
%F_{2}(\vk_1,\vk_2)=\frac{5}{7}+\frac{2}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}(\frac{k_1}{k_2}+\frac{k_2}{k_1}) \\
%G_{2}(\vk_1,\vk_2)=\frac{3}{7}+\frac{4}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}(\frac{k_1}{k_2}+\frac{k_2}{k_1})
%\eea 
%This approach gives a nearly accurate description of $\Lambda$CDM universe \cite{Takahashi:2008to} with slight generalization \cite{Bernardeau:2002rev}:
%\bea
%{\delta}(\vk,\tau) &=&\sum_{n=1}^{\infty} {\delta}^{(n)}(\vk,\tau)=\sum_{n=1}^{\infty}D_1^{n}(\tau)\delta_{n}(\vk) \\
%{\theta}(\vk,\tau)&=&\sum_{n=1}^{\infty}{\theta}^{(n)}(\vk,\tau)\vs
%&=&-\frac{d\,\ln D_1(\tau)}{d\tau}\sum_{n=1}^{\infty}D_1^{n}(\tau)\theta_{n}(\vk) 
%\eea
%Time-dependent second-order density contrast can be expressed as convolution of two linear density fields with kernel $F_2$:
%\be 
%{\delta}^{(2)}(\vk,\tau)=\int \frac{d^{3}\vk_{1}}{(2\pi)^3} F_2(\vk_1,\vk-\vk_1){\delta}^{(1)}(\vk_1,\tau) {\delta}^{(1)}(\vk-\vk_1,\tau) \eql{sorder}
%\ee

@@

\section{Demonstration with N-Body Simulation}\label{sec3}
We test the power of the quadratic estimator using snapshot at $z=0$ data of BigMPDL, one of the MultiDark cosmological simulations \cite{Klypin:2014nov}. Boxsize of BigMDPL is $3.7\,\rm Gpc$. We use dark matter particle data to produce matter inhomogeneities. Galaxy bias \cite{Desjacques:2018rev} should be otherwise considered if galaxy or halo catalog is used.\\
We use nbodykit code to get Fourier density modes \cite{Hand:2018nby} and compare $\delta(\vk_l)$ with $\hat{\delta}(\vk_l)$ for a 3D grid of $\vk_l$ in Fig.~\rf{hist}. $\delta(\vk_l)$ are the long-wavelength modes we get directly from N-body simulation and $\hat{\delta}(\vk_l)$ are the reproduced modes using quadratic estimator \ec{quadest}.\\
\Sfig{hist}{Histogram of number counts of $|\hat{\delta}(\vk_l)/\delta(\vk_l)|$ in Fourier space with different integration range of $k_s$. We can see that in the weakly nonlinear regime, left plot peaks at $1$ which is what we would expect. While for the right plot, when we go higher into nonlinear scale, ratio peak at about $1.3$. This bias comes from the inaccuracy of second-order perturbation theory.}
Ultimate goal of this approach is to reproduce large scale structure in real space. Compare $\delta$ and $\hat{\delta}$ in real space in Fig.~\rf{real} and we can check the robustness of the quadratic estimator for this simulation.
\Sfigg{real}{Comparison between $\delta(\vec{x})$ and $\hat{\delta}(\vec{x})$ and their difference. The 3D grid of data points are unfolded into 2D grids, $7$ columns corresponds to $7$ slices we have. The upper limit of $\vk_s$ is chosen to be $0.15 \, \rm Mpc^{-1}$. We can compute the signal-to-noise ratio as: $S/N=\Var(\delta)/\Var(\hat{\delta}-\delta)\approx 3$.\peikai{comment?}}

>>>>

\date{\today}
\begin{abstract}
\noindent Large scale density modes cannot be measured directly through surveys but we can study them by measuring perturbations on small scales. Cosmological perturbation theory predicts that second-order density inhomogeneities are a convolution of a short- and a long-wavelength mode. This arises physically because small structure structure varies depending on the large scale environment in which it resides. This will induce small scale clustering in real space and an off-diagonal term in two-point statistics in Fourier space. Therefore we can build a quadratic estimator for the large scale field. We demonstrate that this quadratic estimator works pretty well with our N-body simulation test.

\end{abstract}
\maketitle
\section{Introduction}
Measuring the distribution of structures on large scales has been one of the main goals of recent surveys despite it is hard to accomplish directly. To tackle this problem, relations between long- and short-wavelength modes have been frequently discussed recent years \cite{Baldauf:2011fer}\cite{Barreira:2017res}. Method of using small scale position-dependent power spectrum to compute squeezed-limit bispectrum also indicates that we can use small scale perturbation to infer large scale information to some extent \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to fully achieve this goal, we develop in this work a generalized approach of building a quadratic estimator to measure long-wavelength modes indirectly. \\
Starting from second-order standard perturbation theory (SPT) \cite{Goroff:1986sts}\cite{Makino:1992fs}\cite{Jain:1994sop}, we can identify second-order perturbation of a short-wavelength mode as a convolution of a short- and a long-wavelength mode. This can be further related to cosmic microwave background (CMB) lensing case \cite{Hu:2002mr}, where CMB temperature field has a second-order correction due to gravitational field along line of sight. Similarly in our case the short-wavelength mode's nonlinear terms are related to the large scale environment it belongs to.
\\
During CMB lensing quadratic estimator's construction, small scale two-point correlations of CMB temperature modes have off-diagonal terms due to large scale perturbations caused by gravitational lensing. The same statistical feature shows up in our case as well -- off-diagonal terms of small scale correlations are no-longer zero due to large scale modes. Thus we can create a quadratic estimator for long-wavelength modes using the exact same formalism. \\
We begin with a brief review of SPT up to second-order in section \ref{sec1}. Then we build the quadratic estimator and compute the detectability of the estimator in section \ref{sec2}. In section \ref{sec3} we present a concise study of the estimator with data from N-body simulation. Finally we conclude this work and discuss future work related to this in section \ref{sec4}. We use flat $\Lambda$CDM model with Planck Collaboration XVI (2014) \cite{Planck:2014cos} cosmological parameters in this work (same as parameters of the N-body simulation).

@@

In this section we review the SPT approach which describes small scale nonlinearity. Then we compute the off-diagonal term of two short-wavelength modes in the squeezed limit. This non-zero off-diagonal term proportional to a long-wavelength mode will be the key factor of constructing the quadratic estimator. \\

@@

Plugging the perturbative series \ec{pert1} and \ec{pert2} into the Fourier transformed fluid equations \ec{fluid1}-\ec{fluid3}, we can solve second-order density contrast and the result is given by:

@@

Notice \ec{pert1}, \ec{pert2} and \ec{f2} are completely accurate only in Einstein-de Sitter universe. Nonetheless when we actually do related calculation in $\Lambda$CDM universe, the difference is negligible \cite{Takahashi:2008to} thus we still use the expression from \ec{pert1} and \ec{f2} throughout the whole work.\\

@@

@@

\section{Demonstration with N-Body Simulation}\label{sec3}
We test the power of the quadratic estimator using snapshot at $z=0$ data of BigMPDL, one of the MultiDark cosmological simulations \cite{Klypin:2014nov}. Boxsize of BigMDPL is $3.7\,\rm Gpc$. We use dark matter particle data to produce matter inhomogeneities. Galaxy bias \cite{Desjacques:2018rev} should be otherwise considered if galaxy or halo catalog is used.\\
We use nbodykit code to get Fourier density modes \cite{Hand:2018nby} and compare $\delta(\vk_l)$ with $\hat{\delta}(\vk_l)$ for a 3D grid of $\vk_l$ in Fig.~\rf{hist}. $\delta(\vk_l)$ are the long-wavelength modes we get directly from N-body simulation and $\hat{\delta}(\vk_l)$ are the reproduced modes using quadratic estimator \ec{quadest}.\\
\Sfig{hist}{Histogram of number counts of $|\hat{\delta}(\vk_l)/\delta(\vk_l)|$ in Fourier space with different integration range of $k_s$.\peikai{should be more clear what does it count..} We can see that in the weakly nonlinear regime, left plot peaks at $1$ which is what we would expect. While for the right plot, when we go higher into nonlinear scale, ratio peak at about $1.3$. This bias comes from the inaccuracy of second-order perturbation theory.}
Ultimate goal of this approach is to reproduce large scale structure in real space. Compare $\delta$ and $\hat{\delta}$ in real space in Fig.~\rf{real} and we can check the robustness of the quadratic estimator for this simulation.
\Sfigg{real}{Comparison between $\delta(\vec{x})$ and $\hat{\delta}(\vec{x})$ and their difference. The 3D grid of data points are unfolded into 2D grids, $7$ columns corresponds to $7$ slices we have. The upper limit of $\vk_s$ is chosen to be $0.15 \, \rm Mpc^{-1}$. We can compute the signal-to-noise ratio as: $S/N=\Var(\delta)/\Var(\hat{\delta}-\delta)\approx 3$.\peikai{comment?}}
